---
title: "Sorting Hogwarts Hat with Multiple Regression for Harry Potter characters"
author: Shuyuan Han
description: "My final project for Gov 50"
output:
  distill::distill_article: 
    self_contained: false
---

## Research question
As a big fan of the Harry Potter series, I am always wondering how the Sorting Hat actually works. Are the first years randomly assigned to one of the four houses or is there a determined classification algorithm behind? What are the attributes of students in each of the four houses? To find out the science behind the magic, it is a great idea to look at the information of the students (the data, input) and which house they were assigned to (the label, target). In this project, I tried to implement a Sorting Hogwarts Hat with Multiple Regression on R for 1,600 Harry Potter characters based on their personal information.

```{r}
library(tidyverse)
library(ggplot2)
library(infer)
data_students<-read_csv("data.csv")
data_students
```

## Data

The independent variables in this data set will be Best Hand, Arithmancy, Astronomy, Herbology, Defense Against the Dark Arts, Divination, Muggle Studies, Ancient Runes, History of Magic, Transfiguration, Potions, Care of Magical Creatures, Charms, Flying, and the dependent variable, of course, will be Hogwarts House.However, we don't need to look at all the independent variables. It is important to see what variables are useful to group people by. There is no use to use all features as some of them don’t provide an obvious division into groups and can bring only noise. 

```{r}
data_students<-data_students|>
  mutate(Best_hand_Left=ifelse(`Best Hand`=="Left",1,0),
         House_code=case_when(`Hogwarts House`=="Gryffindor"~1,
                              `Hogwarts House`=="Slytherin"~2,
                              `Hogwarts House`=="Ravenclaw"~3,
                              `Hogwarts House`=="Hufflepuff"~4))
cor1<-cor(data_students$House_code,data_students$Best_hand_Left, use = "pairwise")
cor2<-cor(data_students$House_code,data_students$Arithmancy, use = "pairwise")
cor3<-cor(data_students$House_code,data_students$Astronomy, use = "pairwise")
cor4<-cor(data_students$House_code,data_students$Herbology, use = "pairwise")
cor5<-cor(data_students$House_code,data_students$`Defense Against the Dark Arts`, use = "pairwise")
cor6<-cor(data_students$House_code,data_students$Divination, use = "pairwise")
cor7<-cor(data_students$House_code,data_students$`Muggle Studies`, use = "pairwise")
cor8<-cor(data_students$House_code,data_students$`Ancient Runes`, use = "pairwise")
cor9<-cor(data_students$House_code,data_students$`History of Magic`, use = "pairwise")
cor10<-cor(data_students$House_code,data_students$Transfiguration, use = "pairwise")
cor11<-cor(data_students$House_code,data_students$Potions, use = "pairwise")
cor12<-cor(data_students$House_code,data_students$`Care of Magical Creatures`, use = "pairwise")
cor13<-cor(data_students$House_code,data_students$Charms, use = "pairwise")
cor14<-cor(data_students$House_code,data_students$Flying, use = "pairwise")
tibble(independent_variable=c("Best Hand","Arithmancy","Astronomy","Herbology","Defense Against the Dark Arts",	"Divination", "Muggle Studies",	"Ancient Runes","History of Magic",	"Transfiguration","Potions","Care of Magical Creatures", "Charms","Flying"), cor=c(cor1,cor2,cor3,cor4,cor5,cor6,cor7,cor8,cor9,cor10,cor11,cor12,cor13,cor14))


```

From the tibble we narrow our related independent variables to Herbology, History of Magic, Transfiguration, Charms, and Flying. 
 
```{r}
data_students<-data_students|>
  rename(History=`History of Magic`)
regression_data<-lm(House_code~Herbology+History+Transfiguration+Charms+Flying,data = data_students)
regression_data
modelsummary::modelsummary(regression_data,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
gof_map = c("nobs", "r.squared", "adj.r.squared"))
```



