---
title: "Sorting Hogwarts Hat with Multiple Regression for Harry Potter characters"
author: Shuyuan Han
description: "My final project for Gov 50"
output:
  distill::distill_article: 
    self_contained: false
---

## Research question
As a big fan of the Harry Potter series, I am always wondering how the Sorting Hat actually works. Are the first years randomly assigned to one of the four houses or is there a determined classification algorithm behind? What are the attributes of students in each of the four houses? To find out the science behind the magic, it is a great idea to look at the information of the students (the data, input) and which house they were assigned to (the label, target). In this project, I tried to implement a Sorting Hogwarts Hat with Multiple Regression on R for 1,600 Harry Potter characters based on their personal information.

```{r}
library(tidyverse)
library(ggplot2)
library(infer)
data_students<-read_csv("data.csv")
data_students
```

The independent variables in this data set will be Birth month, Best Hand, Arithmancy, Astronomy, Herbology, Defense Against the Dark Arts, Divination, Muggle Studies, Ancient Runes, History of Magic, Transfiguration, Potions, Care of Magical Creatures, Charms, Flying, and the dependent variable, of course, will be Hogwarts House. I used a mutate function here to change the categorical variable, Hogwarts House, to a numerical one based on their performances in the HOUSE CUP competition. With Slytherin ranking first, Gryffindor second, Ravenclaw third, and Hufflepuff last, I assigned 4 to Slytherin, 3 to Gryffindor, 2 to Ravenclaw, and 1 to Hufflepuff as an indicator of overall "excellency" of the house. This study looks at how independent variables, such as birth month, best hand, and performance in each subject affect a student's house assignment. Put it in another way, do these personal traits affect the Sorting Hat's decision to assign a student to different houses in terms of overall "excellency"?

## Does Best Hand Have a Say?
```{r}
data_students<-data_students|>
  mutate(House_code=case_when(`Hogwarts House`=="Hufflepuff"~1,
                              `Hogwarts House`=="Ravenclaw"~2,
                              `Hogwarts House`=="Gryffindor"~3,
                              `Hogwarts House`=="Slytherin"~4))

base_diff<-data_students|>
specify(House_code~`Best Hand`)|>
calculate(stat="diff in means", order=c("Left","Right"))
base_diff

null_diff<-data_students|>
specify(House_code~`Best Hand`)|>
hypothesize(null="independence")|>
generate(reps=1000, type="permute")|>
calculate(stat="diff in means", order=c("Left","Right"))
null_diff

base_p<-null_diff|>
get_p_value(obs_stat = base_diff, direction = "both")
base_p

```
Since the p value is far higher than 0.05 (in fact close to 1), we can retain the null hypothesis. That is, Best Hand does not have a say at all on the house that students were assigned.

```{r}
null_diff|>
visualize()+
shade_p_value(obs_stat = base_diff, direction = "both")
```
## Does Birth Month have a say?
```{r}
library(lubridate)
birthmonth<-month(as.Date(data_students$Birthday, format="%Y/%m/%d"))

data_students<-data_students|>
  mutate(Birthmonth=birthmonth)

ggplot(data_students, mapping=aes(x=Birthmonth, fill=`Hogwarts House`))+
  geom_bar()+
  labs(x="Birth Month",
       y="Number of Students")
  
```
From the graph above, we see no evidence of one's birth month deciding his or her House assignment.For each house, students are basically evenly distributed into the 12 months.

## How is Course performance related to House assignment?

```{r}

cor1<-cor(data_students$House_code,data_students$Arithmancy, use = "pairwise")
cor2<-cor(data_students$House_code,data_students$Astronomy, use = "pairwise")
cor3<-cor(data_students$House_code,data_students$Herbology, use = "pairwise")
cor4<-cor(data_students$House_code,data_students$`Defense Against the Dark Arts`, use = "pairwise")
cor5<-cor(data_students$House_code,data_students$Divination, use = "pairwise")
cor6<-cor(data_students$House_code,data_students$`Muggle Studies`, use = "pairwise")
cor7<-cor(data_students$House_code,data_students$`Ancient Runes`, use = "pairwise")
cor8<-cor(data_students$House_code,data_students$`History of Magic`, use = "pairwise")
cor9<-cor(data_students$House_code,data_students$Transfiguration, use = "pairwise")
cor10<-cor(data_students$House_code,data_students$Potions, use = "pairwise")
cor11<-cor(data_students$House_code,data_students$`Care of Magical Creatures`, use = "pairwise")
cor12<-cor(data_students$House_code,data_students$Charms, use = "pairwise")
cor13<-cor(data_students$House_code,data_students$Flying, use = "pairwise")
tibble1<-tibble(independent_variable=c("Arithmancy","Astronomy","Herbology","Defense Against the Dark Arts",	"Divination", "Muggle Studies",	"Ancient Runes","History of Magic",	"Transfiguration","Potions","Care of Magical Creatures", "Charms","Flying"), cor=c(cor1,cor2,cor3,cor4,cor5,cor6,cor7,cor8,cor9,cor10,cor11,cor12,cor13))

knitr::kable (tibble1)

library(GGally)
GGally::ggcorr(data_students,
               methods=c("everything","pearson"),
               label = TRUE,
               label_alpha = TRUE)
```

To build a multiple regression model,we don't need to use all the independent variables. It is important to see what variables are useful to group people by. There is no use to use all features as some of them don’t provide an obvious division into groups and can bring only noise. From the table above we can narrow our related independent variables to Herbology, Divination, Defense against the Dark Arts, and Charms that have a relatively strong correlation with the House_code.
 
```{r}
library(tidyverse)
data_students<-data_students|>
  rename(Dark= `Defense Against the Dark Arts`)
regression_data<-lm(House_code~Dark+Herbology+Divination+Charms,data = data_students)
regression_data
modelsummary::modelsummary(regression_data,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

Holding other variables constant, a one unit change in the independent variable Herbology would result in a change of negative 0.145 in the House_code (negative correlation), a one unit change in Divination would result in a change of negative 0.014 in the House_code (negative correlation), a one unit change in Defense Against the Dark Arts would result in a change of 0.1 in the House_code (positive correlation), and a one unit change in Charms would result in a change of negative 0.015 in the House_code (negative correlation).An estimated coefficient is statistically significant when its p value is less than alpha. Therefore, at the 0.05 level, Herbology, Defense against the Dark Arts, and Charms are all statistically significant, but Divination is not. 
 
## Conclusion
According to the discussion above, the birth month of a student and his or her best hand have barely nothing to do with the Sorting Hat's decision. As for performance in each of the main courses, a student's performance in Herbology and Divination are both negatively correlated with the overall "excellency" of the house he or she is assigned to, and a student's performance in Defense Against the Dark Arts is positively correlated with the overall "excellency" of the house he or she is assigned to. And these are statistically significant. This makes sense in that Defense Against the Dark Arts is assumed to be the most important skill in a world where Dark Art threatens the lives of almost everyone. An excellent wizard should be well equipped with knowledge in this field. Herbology and Divination, however, may make students a good scholar in academia, but not necessarily an excellent wizard. This study, however, only looks at the birth month, best hand, and performance in main courses. There might be many other variables that affect the Sorting Hat's decision, for example, the students' family background, characteristics, hometown, and so on. Further study is therefore needed to look at those variables.
 